Concurrent Priority Queues

### 2014-03-24 presentation/presentation.tex
Lockfree section

\begin{frame}{\nameref{sec:locking}}
\begin{itemize}
\item Naive PQ parallelization: single global lock \textrightarrow ~ sequential bottleneck
\item A first improvement: fine-grained locking using a lock per node
\item \fullcite{hunt1996efficient}
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:locking}}
\framesubtitle{Innovations}

@@

\begin{frame}{\nameref{sec:locking}}
\framesubtitle{Limitations}

@@

\begin{frame}{\nameref{sec:lockfree}}

>>>>

\begin{frame}{\nameref{sec:locking}}
\framesubtitle{\citeauthor{hunt1996efficient}}

\begin{itemize}
\item Naive PQ parallelization: single global lock \textrightarrow ~ sequential bottleneck
\item A first improvement: fine-grained locking using a lock per node
\item \fullcite{hunt1996efficient}
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:locking}}
\framesubtitle{\citeauthor{hunt1996efficient}: Innovations}

@@

\begin{frame}{\nameref{sec:locking}}
\framesubtitle{\citeauthor{hunt1996efficient}: Limitations}

@@

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{SkipLists}

\begin{itemize}
\item Modern concurrent PQs are mostly based on \citeauthor{pugh1990skip}'s SkipList
\item Probabilistic ordered search structure, insertions and deletions in expected $O(\log n)$ time
\item No reorganizations
\item Simple implementation
\item Excellent disjoint-access properties
\end{itemize}
\end{frame}

\begin{frame}[fragile]{\nameref{sec:lockfree}}
\framesubtitle{SkipLists}

\begin{itemize}
\item Collection of linked lists with corresponding levels
\item Lowest list contains all items, higher lists are shortcuts
\item \lstinline|Insert| chooses a \lstinline|level| according to geometric distribution
\end{itemize}

\begin{lstlisting}
struct slist_t {
  size_t max_level;
  node_t head[max_level];
};
struct node_t {
  key_t key;
  value_t value;
  size_t level;
  node_t *next[level];
};
\end{lstlisting}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{shavit2000skiplist}}

\begin{itemize}
\item First SkipList-based PQ
\item \fullcite{shavit2000skiplist}
\item Initially lock-based, lock-free variant published in 2008
\item \fullcite{herlihy2012art}
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{shavit2000skiplist}}

\begin{itemize}
\item Items are considered in the list once inserted on bottom level
\item Again, insertions bottom-up and deletions top-down
\item Deletions are split
    \begin{itemize}
    \item Logical deletion sets a \lstinline|deleted| flag
    \item Physical deletion performs actual pointer manipulations
    \end{itemize}
\item \lstinline|DeleteMin| attempts to logically delete the head node.
     On success: delete physically \& return node. Otherwise, continue with
     next node.
\item \lstinline|Insert| is equivalent to the SL insertion
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{shavit2000skiplist}}

\begin{itemize}
\item
\end{itemize}
\end{frame}
